---
title: "Munging Students & AP Data"
output: html_notebook
---

# double check we are getting the best scores, filter out NA students, clean names
```{r}
ap_data <- ap_results_codes %>%
  group_by(StudentNumberDelta,
           ExamCode,
           ExamName,
           AdminYear,
           AcademicYear,
           TookTest) %>%
  summarize(AP_Result = max(AP_Result, na.rm=TRUE)) %>%
  distinct() %>%
  #filter(!is.na(ap_file_path)) %>%
  janitor::clean_names() %>%
  ungroup()
```

# Collect all distinct student numbers from all years
```{r}
stu_scaffold <-
  ap_data %>% ungroup %>%
  select(student_number = student_number_delta,
         academic_year) %>%
  bind_rows(students_df %>%
              ungroup() %>%
              select(student_number, academic_year) )%>%
  arrange(student_number, academic_year) %>%
  distinct()
```

# Scaffold admin years
```{r}
admin_year_scaffold <- tibble(
  academic_year = sprintf("%s-%s", 2009:2021, 2010:2022)) %>%
  mutate(admin_year = as.integer(str_extract(academic_year, regex("\\d{2}$")))-9)
```

# list of IB Schools
```{r}
ib_exception_schools <- c("Donna", # just works IB, offer very few APS
                          "South Flores", # not trying to hit AP goal
                          "Brownsville",  # include if it helps this year. Should be included in 21-22
                          "Frontier",      # try hit AP Goal in 20-21
                          "McAllen"
)


```

# use studnet scaffold to ensure we are tracking all studunts, and not just
# students who tested
```{r}
ap_full <- stu_scaffold %>%
  left_join(students_df, by = c("student_number", "academic_year")) %>%
  left_join(ap_data, by = c("student_number" = "student_number_delta",
                            "academic_year" = "academic_year")) %>%
  select(student_number,
         academic_year,
         admin_year,
         school_name,
         region,
         grade_level_id,
         took_test,
         ap_result,
         exam_name,
         enrollment_status,
         graduate,
         race,
         sped,
         at_risk,
         ecd,
         lep
  ) %>%
  # last check on max exam results
  mutate(ap_result = as.integer(ap_result)) %>%
  group_by(student_number
           ,academic_year
           ,admin_year
           ,school_name
           ,region
           ,grade_level_id
           ,enrollment_status
           ,graduate
           ,took_test
           ,exam_name
           ,race
           ,sped
           ,at_risk
           ,ecd
           ,lep) %>%
  summarise(ap_result = max(ap_result)) %>%
  distinct() %>%

  # this next set of steps ensures that every student who didn't take the test
  # has an admin_year, since that data came from the College Board data
  left_join(admin_year_scaffold, by = "academic_year") %>%
  select(-admin_year.x) %>%
  rename(admin_year = admin_year.y) %>%
  mutate(took_test = if_else(is.na(took_test), 0, 1),

         #this school name renaming seems unneccesary
         school_name = if_else(is.na(school_name), "Placeholder", school_name)) %>%

  # need to indicate if score sim from a student at IB focused campuses
  mutate(ib_school = school_name %in% ib_exception_schools)
  # mutate(ib_school = case_when(academic_year <= "2020-2021" & school_name %in% ib_exception_schools_20_21 ~ TRUE,
  #                              school_name %in% ib_exception_schools_21_22 ~ TRUE,
  #                              TRUE ~ FALSE))
```

# Exam group scaffold -----

# order matters here since it needs to match the grades in which these are
# generally taught
```{r}
exams_vector <- c("Spanish Language and Culture",
                  "World History",
                  "Human Geography",
                  "Physics 1",
                  "United States Government and Politics",
                  "English Language and Composition",
                  "United States History",
                  "Biology",
                  "English Literature and Composition",
                  "Microeconomics",
                  "Calculus AB",
                  "Chemistry",
                  "Spanish Literature and Culture",
                  "Statistics",
                  "Environmental Science")

# so order matters here too
grade_level_vector <- c(11,10,9,12,12,11,11,12,12,12,12,12,12,12,12)


exams_df <- tibble(exam_name = exams_vector,
                   grade_level_id = grade_level_vector) %>%
  mutate(concat =  paste(exam_name, grade_level_id),
         ones = 1, # I kinda hate this
         ap_group = case_when(
           str_detect(exam_name, "Bio|Chem|Environ|Physics") ~ "Science",
           str_detect(exam_name, "World|Human|Microeconomics|United States") ~ "Humanities",
           str_detect(exam_name, "Calculus AB|Statistics") ~ "Math",
           str_detect(exam_name, "Spanish") ~ "Spanish",
           str_detect(exam_name, "English") ~ "English",
           TRUE ~ "Other"
         ))
```

```{r}
ap_df <- ap_full %>%
  left_join(exams_df %>% select(exam_name, ap_group),
            by = c("exam_name" = "exam_name")) %>%
  mutate(ap_group = if_else(is.na(ap_group), "Other", ap_group),
         ap_group = if_else(!is.na(exam_name), ap_group, as.character(NA))) %>%
  ungroup() %>%
  arrange(student_number, academic_year, exam_name) %>%
  group_by(student_number) %>%
  mutate(passed = ap_result >= 3,
         cum_passed = cumsum(replace_na(as.integer(passed), 0))) %>%
  ungroup() %>%
  filter(!is.na(student_number))
```

